Pre/Post-Trikafta

### 2022-01-19 Christine Chan/Pre-post Trikafta/pre_post_trikafta.Rmd
Update pre_post_trikafta.Rmd
function to get closest visits to start date

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lubridate)
library(readxl)
library(PAutilities)
library(AGD)
library(rspiro)
library(anthro)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta/Data_Raw")
```

```{r warning=FALSE,results='asis'}
############pre/post trikaftka
# c. chan
##############################

#cgm data
# cgm_c <- read.csv("REDCap Upload.csv")
cgm_c <- read.csv("./Data From CHCO/REDCap Upload_chco.csv")
cgm_c$date_cgm <- as.Date(cgm_c$date_cgm_placement, "%Y-%m-%d")

cgm_m <- read.csv("./Data From Montana/CGM_Montana/REDCap Upload_mt.csv")
cgm_m$date_cgm <- as.Date(cgm_m$date_cgm_placement, "%m/%d/%Y")

cgm_w <- read.csv("./Data From Washington/REDCap Upload_Washington.csv")
cgm_w$date_cgm <- as.Date(cgm_w$date_cgm_placement, "%Y-%m-%d")

cgm <- rbind(cgm_c, cgm_m,cgm_w )

`%!in%` = Negate(`%in%`)
reg0 <- read.csv("./Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
#need to get rid of those two
reg <- subset(reg0, reg0$MRN %!in% c("1714343", "787372") )

reg$Earliest.Trikafta.Date.in.PortCF <- ifelse(reg$MRN =="1053930", "12/1/2019", 
                                               ifelse(reg$MRN =="1695512", "12/12/2019", 
                                                      ifelse(reg$MRN =="815274", "12/13/2019", 
                                                             ifelse(reg$MRN =="1078408",   "3/9/2020",
                                                                    ifelse(reg$MRN =="899717", "2/14/2020", as.character(reg$Earliest.Trikafta.Date.in.PortCF))))))

a1c <- read.csv("./Data From CHCO/Copy of A1C and OGTTs Before and After Trikafta_cc modified 8.25.20.csv")

weight <- read.csv("./Data From CHCO/Registry/weightdata.csv")

weight$Earliest.Trikafta.Date.in.PCF <- ifelse(weight$MRN =="1053930", "12/1/2019", 
                                               ifelse(weight$MRN =="1695512", "12/12/2019", 
                                                      ifelse(weight$MRN =="815274", "12/13/2019", 
                                                             ifelse(weight$MRN =="1078408",   "3/9/2020",
                                                                    ifelse(weight$MRN =="899717", "2/14/2020", as.character(weight$Earliest.Trikafta.Date.in.PCF))))))

#montana data
#total population

library("readxl")

reg_montana <- read.csv("./Data From Montana/Copy of Paper patient population glycemic (+hofer).csv")

demo_adult <- read.csv("./Data From Montana/demo_adults.csv")
demo_kids <- read.csv("./Data From Montana/demo_kids.csv")

reg_montana$age <- ifelse(reg_montana$CFF.ID %in% demo_adult$CFF.ID, "Adult", 
                          ifelse(reg_montana$CFF.ID %in% demo_kids$CFF.ID, "Kids", "Check"))

reg_montana$trik <- as.Date(reg_montana$Trikafta.start.date,"%m/%d/%Y")

reg_montana$test <- as.Date(reg_montana$Date.of.Encounter,  "%m/%d/%Y")
reg_montana$prepost <- ifelse(reg_montana$trik > reg_montana$test, "before", "after")

reg_montana$diff <- reg_montana$trik-reg_montana$test 
reg_montana2 <- subset(reg_montana, abs(diff) <= 365)

reg_montana3 <- reg_montana2 %>% group_by(CFF.ID, prepost) %>% count()
reg_montana4 <- reg_montana3 %>% group_by(CFF.ID) %>% count()

#weight/fev

pft_adult <- read.csv("./Data From Montana/Copy of PFTs growth adult (+hofer).csv")
pft_kids <- read.csv("./Data From Montana/PFTs kids.csv")

pft_montana <- rbind(pft_adult, pft_kids)

pft_montana_1 <- subset(pft_montana, pft_montana$CFF.ID %in% reg_montana$CFF.ID)

#age at pft
pft_montana_1$pft <- as.Date(pft_montana_1$Date.of.Encounter,  "%m/%d/%Y")

reg_montana_first <- reg_montana %>% group_by(CFF.ID) %>% filter(row_number()==1)

reg_montana_first$trik <- ifelse(reg_montana_first$CFF.ID == "1044621", "2019-12-02", 
                                 ifelse(reg_montana_first$CFF.ID == "1008380", "2020-1-1", as.character(reg_montana_first$trik)))

pft_montana_2 <- merge(pft_montana_1, reg_montana_first[,c("CFF.ID" ,"Date.Of.Birth..mm.dd.yyyy.", "trik")])

#need pfts within a year of trik

pft_montana_2$diff <-as.numeric(as.Date(pft_montana_2$pft) - as.Date(pft_montana_2$trik))
pft_montana_3 <- subset(pft_montana_2, abs(diff) <= 365 )
pft_montana_3$prepost <- ifelse(pft_montana_3$trik > pft_montana_3$pft, "before", "after")
pft_montana_4 <- subset(pft_montana_3, pft_montana_3$prepost == "before")
pft_montana_5 <- pft_montana_4 %>% group_by(CFF.ID) %>% arrange(diff) %>% filter(row_number()==1)

pft_montana_5$DOB <- as.Date(pft_montana_5$Date.Of.Birth..mm.dd.yyyy.,  "%m/%d/%Y")

pft_montana_5$age <- as.numeric(pft_montana_5$pft-pft_montana_5$DOB)/365.25

#need sex
demo_montana <- rbind(demo_kids[,c("CFF.ID","Gender")], demo_adult[,c("CFF.ID","Gender")])

analytic_montana <- merge(pft_montana_5, demo_montana ,by = "CFF.ID", all.x=T)

#need cfrd start date - don't have dates

#gtube status

gt_adult <- read.csv("./Data From Montana/gtubes_adult.csv")
gt_kids <- read.csv("./Data From Montana/GT users kids.csv")

gt_montana <- rbind(gt_adult, gt_kids)

gt_montana_1 <- merge(gt_montana, reg_montana_first[,c("CFF.ID" , "trik")])
gt_montana_1$gt <- as.Date(gt_montana_1$Date.of.Encounter,  "%m/%d/%Y")
gt_montana_1$diff <-as.numeric(as.Date(gt_montana_1$gt) - as.Date(gt_montana_1$trik))
gt_montana_2 <- subset(gt_montana_1, abs(diff) <= 365 )
analytic_montana$gtube <- ifelse(analytic_montana$CFF.ID %in% gt_montana_2$CFF.ID, "Yes", "No")

#Insufficent

is_adult <- read.csv("./Data From Montana/pancreatic insufficient adult.csv")
is_kids <- read.csv("./Data From Montana/pancreatic insufficient kids.csv")

analytic_montana$is <- ifelse(analytic_montana$CFF.ID %in% c(is_adult$CFF.ID, is_kids$CFF.ID), "Insufficient", "Not")

#genotypes

g_adult <- read.csv("./Data From Montana/genotypes_adult.csv")
g_kids <- read.csv("./Data From Montana/genotypes kids.csv")[,1:3]

geno_montana <- rbind(g_adult,g_kids)

geno_montana$Genotypes1 <- geno_montana$Name.of.the.first.mutation

geno_montana$Genotypes2 <- geno_montana$Name.of.the.second.mutation

geno_montana$class2 <- ifelse(geno_montana$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
                              ifelse(geno_montana$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
                                     "II", 
                                     ifelse(geno_montana$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
                                            ifelse(geno_montana$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
                                                   "IV", 
                                                   ifelse(geno_montana$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))

geno_montana$severity2 <- ifelse(geno_montana$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
                                                                '2183delAA->G', '2184insA', 'S434X',
                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
                                 "Severe", 
                                 ifelse(geno_montana$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,

                                        ifelse(geno_montana$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
                                                                                "R1066H", "P205S"), 
                                               "Unk","")))

geno_montana$class1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                             '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                             '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                             '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
                              ifelse(geno_montana$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
                                     "II", 
                                     ifelse(geno_montana$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
                                            ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
                                                   "IV", 
                                                   ifelse(geno_montana$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                                         'P574H', '3272-26A->G5'), "V" , "")))))   
geno_montana$severity1 <- ifelse(geno_montana$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                                '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                                '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                                '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
                                                                'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
                                                                '2183delAA->G', '2184insA', 'S434X',
                                                                '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
                                                                'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
                                                                "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
                                 "Severe", 
                                 ifelse(geno_montana$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
                                                                       'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                       'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
                                                                       'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
                                                                       "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,

                                        ifelse(geno_montana$Genotypes1 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
                                                                                'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
                                                                                'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
                                                                                'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
                                                                                "R1066H", "P205S"), 
                                               "Unk","")))

geno_montana$genoRisk <- ifelse(geno_montana$severity1 == "Mild" | geno_montana$severity2 == "Mild", "Low", 
                                ifelse(geno_montana$severity1 == "Severe" & geno_montana$severity2 == "Severe", "High", ""))

analytic_montana$genotype <- ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "High", ]$CFF.ID, "High", 
                                    ifelse(analytic_montana$CFF.ID %in% geno_montana[geno_montana$genoRisk == "Low", ]$CFF.ID, "Low",   ""   ))

#washingotn

reg_washington <- read.csv("./Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")
reg_wash <- read.csv("./Data From Washington/Copy of CFRD and trikafta Trikafta_on_Annual_Review-13-Apr-2021-0815.csv")

analytic_wash <- reg_washington %>% group_by(CFF.ID) %>% filter(row_number()==1) %>% filter(CFF.ID != "1499770")
analytic_wash$trik <- as.Date(analytic_wash$Approximate.date.patient.started.taking.this.drug,  "%m/%d/%Y")

#aic 
a1c_wash <- read.csv("./Data From Washington/A1C_and_OGTT.csv")

temp_wash <- reg_wash[,c("CFF.ID", "Approximate.date.patient.started.taking.this.drug")]
temp_wash1 <- temp_wash %>% group_by(CFF.ID) %>% filter(row_number()==1)

a1c_wash1 <- merge(a1c_wash, temp_wash1, by = c("CFF.ID"))
a1c_wash1$pft <- as.Date(a1c_wash1$Date.of.Encounter,  "%m/%d/%Y")
a1c_wash1$trik <- as.Date(a1c_wash1$Approximate.date.patient.started.taking.this.drug,  "%m/%d/%Y")
a1c_wash1$diff <-as.numeric(a1c_wash1$pft - a1c_wash1$trik)
a1c_wash2<- subset(a1c_wash1, abs(diff) <= 365 )

#bmi
bmi_wash <- read.csv("./Data From Washington/BMIandWeight.csv")
bmi_wash1 <- merge(bmi_wash, analytic_wash, by = "CFF.ID")
bmi_wash1$Date <- as.Date(bmi_wash1$Date.of.Encounter.x,  "%m/%d/%Y")
bmi_wash1$diff <-as.numeric(bmi_wash1$Date - bmi_wash1$trik)

bmi_wash2 <- subset(bmi_wash1, abs(diff) <= 365 )
bmi_wash2$prepost <- ifelse(bmi_wash2$trik > bmi_wash2$Date, "before", "after")

#need age/bmi from first visit
bmi_wash3 <- bmi_wash2 %>% group_by(CFF.ID) %>% filter(prepost == "before") %>% arrange(diff) %>%  filter(row_number()==1)
bmi_wash3$first_encounter <- bmi_wash3$Date.of.Encounter.x

analytic_wash <- merge(analytic_wash,bmi_wash3[,c("CFF.ID","BMI.Percentile" ,"Encounter.Age"  ,"Patients.height","BMI.Value","first_encounter")], by = "CFF.ID")

#sex
demo_wash <- read.csv("./Data From Washington/Demographics.csv")

analytic_wash$Sex <- ifelse(analytic_wash$CFF.ID %in% demo_wash[demo_wash$Gender== "Female",]$CFF.ID, "F", 
                            ifelse(analytic_wash$CFF.ID %in% demo_wash[demo_wash$Gender== "Male",]$CFF.ID, "M",  "Check"))

#Insufficent
is_wash <- read.csv("./Data From Washington/Pancreatic_Insufficient.csv")

analytic_wash$is <- ifelse(analytic_wash$CFF.ID %in% is_wash$CFF.ID, "Insufficient", "Not")

#gtube

gt_wash <- read.csv("./Data From Washington/G-tube.csv")

analytic_wash$gt <- ifelse(analytic_wash$CFF.ID %in% gt_wash$CFF.ID, "GT", "Not")

#genotypes

geno_wash <- read.csv("./Data From Washington/Genotypes.csv")

geno_wash$Genotypes1 <- geno_wash$Name.of.the.first.mutation

geno_wash$Genotypes2 <- geno_wash$Name.of.the.second.mutation

#
geno_wash$class2 <- ifelse(geno_wash$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
                           ifelse(geno_wash$Genotypes2 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
                                  "II", 
                                  ifelse(geno_wash$Genotypes2 %in% c('G551D', 'G551S', 'S549N'), "III", 
                                         ifelse(geno_wash$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
                                                "IV", 
                                                ifelse(geno_wash$Genotypes2 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))

geno_wash$severity2 <- ifelse(geno_wash$Genotypes2 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
                                                          '2183delAA->G', '2184insA', 'S434X',
                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
                              "Severe", 
                              ifelse(geno_wash$Genotypes2 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,

                                     ifelse(geno_wash$Genotypes2 %in% c(  '1262insA', 'I506T', '3363delGT' , 'Q1209P', '3349insT', 'G178R', 'L467P', '317insC', 'I336K',	
                                                                          'ex14a', '3500-2A->G', 'c.3407_3422del16', '3349insT', '1248+1G->A', 'M470V', '1812-1G->A',
                                                                          'del X22-24', 'Q237H', 'Q237H', 'W57G', '1078delA', 'G178R', '2957delT', 'G576A',
                                                                          'other', 'Other', 'Unknown', 'unknown', 'CFTRdele1', '1811+1.6kb A->', 'UNK', "I336K", "M470V", "S1235R", "M470V", "F508C",
                                                                          "R1066H", "P205S"), 
                                            "Unk","")))

geno_wash$class1 <- ifelse(geno_wash$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                       '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                       '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                       '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G'), "I", 
                           ifelse(geno_wash$Genotypes1 %in% c('F508', 'F508del', 'F508del ', 'G85E', 'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E'), 
                                  "II", 
                                  ifelse(geno_wash$Genotypes1 %in% c('G551D', 'G551S', 'S549N'), "III", 
                                         ifelse(geno_wash$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T'), 
                                                "IV", 
                                                ifelse(geno_wash$Genotypes1 %in% c('A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                                   'P574H', '3272-26A->G5'), "V" , "")))))   
geno_wash$severity1 <- ifelse(geno_wash$Genotypes1 %in% c('1717-1G->A', '1717-1G-A',  'G542X', 'Q493X', 'R1162X', 'R553X', 'R553x', 'W1089X', 'W1282X', '1078delT', 'R75X',
                                                          '3659delC', '621+1G->T', '621+1G>T', '394delTT', '3120 + 1g->A', '1154InsTC',  '1154insTC', '1213delT',
                                                          '1259insA', '1288insTA', '3791delC', 'E60X', 'K710X', '2184delA', 'CFTRdele2,3', '663delT', 'Glu528',
                                                          '1461ins4', '306insA', 'R709X', 'CFTRdele22-24', '2711delT', '2183AA- >G','F508', 'F508del', 'F508del ', 'G85E', 
                                                          'I507', 'I 507', 'N1303K', 'A559T', 'R560T', 'A561E','G551D', 'G551S', 'S549N','1898+1G->A', '2143delT', '2183AA-G',
                                                          '2183delAA->G', '2184insA', 'S434X',
                                                          '2585delT', '3120+1G->A',  '3905insT',  '406-1G->A', '712-1G->T', 'L1245X', 'Q2X', 'R1158X',
                                                          'R851X', 'V520F', 'L1077P', 'S489X', 'F311del', 'W846X', 'M1101K', '2622+1G->A', 'S945L', 
                                                          "Y1092X", "R1066C", "G178R", "2957delT", "CFTRdele1", 'W1204X', "q493x", "W57G", "1585-8G>A", "1717-8G->A","G1244E"), 
                              "Severe", 
                              ifelse(geno_wash$Genotypes1 %in% c('R334W', 'R347P', 'R117H', 'R117C', 'P67L', 'L206W', 'I206w', 'D614G', 'R347H', 'D1152H', '3849+10kbC>T', "D110H", "R1070W",
                                                                 'A455E', '2789+5G->A', '3849+10CT', '3849+10kbC->T', '711+3A->G', '1898+5G->T',
                                                                 'P574H', '3272-26A->G5','3849G->A', 'H199Y', 'H199y',  'S492F', 'S531P',
                                                                 'L1335P', 'R352Q', "D1270N", "2789+2insA", "3849+10kbc->T", "Q98R", "Phe1052Val", 
                                                                 "F1052V", "Tyr1014Cys", "Y1014C", "T338I", "5T", "c.2249C>T", "P750L"), "Mild" ,

>>>>

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lubridate)
library(readxl)
library(PAutilities)
library(AGD)
library(rspiro)
library(anthro)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Prepost Triakfta")
```

```{r code from maxie,eval=FALSE}
# Source Maxie's code, and write final ID lists
source("~/GitHub/CF/Christine Chan/Pre-post Trikafta/p001+cmw_UPDATED.R")
ids = data.frame(unique(analytic_cm$CFF.ID))
colnames(ids) = "CFF.ID"
write.csv(ids,file = "./Data_Cleaned/final_ids.csv",row.names = F)
```

```{r closest date function}
closest_dates = function(df,start = "Start",date = "Date"){
  # Start and date columns must be formated for subtraction (ideally with lubridate)
  df$Timepoint = as.numeric(df[,date] - df[,start])
  # Remove visits on start date or > 1 year out
  df = df %>% filter(Timepoint !=0 & abs(Timepoint) <= 365) %>% 
    # Get closest visit before and closest after
    group_by(CFF.ID,sign(Timepoint)) %>%
    slice_min(Timepoint * sign(Timepoint),n=1) %>% 
    # Format
    ungroup %>% select(-Timepoint) %>% rename(Timepoint = `sign(Timepoint)`)
  # Timepoint as factor
  df$Timepoint = factor(df$Timepoint,labels = c("Before","After"))
  # Delete unnecessary columns
  df = df %>% select(-all_of(c(start,date)))
  # Return
  df = data.frame(df)
  return(df)
} 
```

# Table 1: Demographics at Trikafta Start

```{r}
# Start dates
chco_demographics = read.csv("./Data_Cleaned/chco_trikafta_dates.csv",na.strings = "")
# Fix select dates
chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1053930"] = "12/1/2019"
chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1695512"] = "12/12/2019"
chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="815274"] = "12/13/2019"
chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="1078408"] = "3/9/2020"
chco_demographics$Earliest.Trikafta.Date.in.PortCF[chco_demographics$MRN =="899717"] = "2/14/2020"
chco_demographics$CFF.ID = as.character(chco_demographics$CFF.ID)
# Diabetes and with start date only
chco_demographics = chco_demographics %>% 
  filter(!is.na(Earliest.Trikafta.Date.in.PortCF),CFRD.yes.1 == 1)
chco_demographics$Earliest.Trikafta.Date.in.PortCF = 
  mdy(chco_demographics$Earliest.Trikafta.Date.in.PortCF)
chco_dates = chco_demographics %>% rename(Start = Earliest.Trikafta.Date.in.PortCF) %>%
  select(CFF.ID,MRN,Start)
# A1C and OGTT
chco_a1c = read.csv("./Data_Cleaned/chco_a1c_ogtt.csv")
## Format
chco_a1c$CFF.ID = as.character(chco_a1c$CFF.ID)
chco_a1c$Date = mdy(chco_a1c$Date)
chco_a1c$Diabetes = NULL
## Only those w/CFRD and known start
chco_a1c = left_join(chco_dates,chco_a1c,by = "CFF.ID")
## Get closest before and closest after dates
chco_a1c = closest_dates(chco_a1c)
# PFTs
chco_pft = read.csv("./Data_Cleaned/chco_bmi_pft.csv")
chco_pft$Date = mdy(chco_pft$Date)
chco_pft = left_join(chco_dates,chco_pft,by = "MRN")
chco_pft = closest_dates(chco_pft)
# CGM
chco_cgm = read.csv("./Data_Cleaned/chco_cgm.csv")
chco_cgm$Date = date(mdy_hm(chco_cgm$date_cgm_placement))
chco_cgm$CFF.ID = sub("_.*","",chco_cgm$subject_id)
# chco_cgm$subject_id = NULL
chco_cgm = left_join(chco_dates,chco_cgm,by = "CFF.ID")
chco_cgm = closest_dates(chco_cgm)
# Merge
chco = full_join(chco_a1c,chco_pft,by = c("CFF.ID", "MRN", "Timepoint"))
chco = full_join(t,chco_cgm,by = c("CFF.ID", "MRN", "Timepoint"))

# Montana
montana = read.csv("./Data_Raw/Data From Montana/Copy of Paper patient population glycemic (+hofer).csv",na.strings = "")
demo_adult = read.csv("./Data_Raw/Data From Montana/demo_adults.csv",na.strings = "")
demo_kids = read.csv("./Data_Raw/Data From Montana/demo_kids.csv",na.strings = "")
demo_kids$Date.Of.Birth..mm.dd.yyyy. = NULL
demo_kids = demo_kids[,colnames(demo_adult)]
demo = rbind(demo_adult,demo_kids)
# Get within a year of Trikafta start
montana$time_diff = mdy(montana$Trikafta.start.date) - mdy(montana$Date.of.Encounter)
montana = montana %>% filter(abs(montana$time_diff)<=365)
# Add demographics
montana = left_join(montana,demo,by = "CFF.ID")
# PFTs
pft_adult <- read.csv("./Data_Raw/Data From Montana/Copy of PFTs growth adult (+hofer).csv")
pft_kids <- read.csv("./Data_Raw/Data From Montana/PFTs kids.csv")
pft_adult = pft_adult[,colnames(pft_kids)]
pfts = rbind(pft_adult,pft_kids)
montana = full_join(montana,pfts,by = c("CFF.ID", "Date.of.Encounter"))
```
